Modeling the impact of IT- and AI-related emissions in a university setting

### 2025-08-03 doc/paper.tex
Checkpoint

\documentclass[11pt]{article}
\usepackage[margin=0.7in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{libertine}
\renewcommand{\sfdefault}{fvs} % Use as sans-serif
\renewcommand{\ttdefault}{fvm} % Use as monospace
\renewcommand{\ttfamily}{\small\fontfamily{\ttdefault}\selectfont} % Slightly smaller monospace font
\newcommand{\assumption}[1]{{#1}}
\newcommand{\eg}{{\em e.g.}}
%\usepackage{helvet} % Helvetica font for sans-serif
%\usepackage{courier} % Courier font for monospace
\usepackage{hyperref}
\hypersetup{
    hidelinks
}
\usepackage[style=numeric]{biblatex}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{titling}

@@

Information technology accounts for over 5\% of global electricity consumption and at least 1.7\% of global carbon emissions~\cite{wb:itu:ict}. At the University of Zürich (UZH), which with over 38,000 students and staff is the largest Swiss university, IT in 2024 accounted for about 1.5 Mt CO$_2$e (megatons of carbon dioxide equivalent) of greenhouse gas (GHG) emissions, approximately 6.3\% of total university emissions in 2024~\cite{uzh:sustainability:report}. The cited report takes into account university-owned workstations, servers, and audiovisual equipment, but does include other IT-related activities such as networking, cloud computing, or the growing use of artificial intelligence, and in particular the adoption of generative AI and large language models (LLMs).

@@

\subsubsection*{Devices}

@@

University-owned devices are in scope. We have accurate numbers for staff workstations, classroom computers, network access points, and audiovisual devices managed by the central IT department. We can only estimate other devices, such as computers managed by individual research groups or networking equipment. As before, we include these devices' embodied emissions in the model.

@@

Due to a lack of data we do not attempt to model emissions from application-level cloud services such as Microsoft 365 or Google Workspace. This omission leads to undercounting total IT energy use but to overestimating the potential relative impact of AI.

UZH faculties do make use of shared off-campus supercomputing facilities, primarily the Swiss National Supercomputing Center, CSCS~\cite{cscs}. We estimate the impact of these workloads based on available gross peak load numbers. Due to lack of data, we do not include the embodied emissions of these facilities.

@@

We estimate inference-time emissions of current large language models based on the literature. It is harder to come by good estimates of training-time emissions and embodied emissions of the hardware, but we attempt to do so nonetheless in order to not risk undercounting the impact of LLMs.

@@

\subsection{Calculation approach}

\subsubsection*{Devices}

We model campus device emissions bottom-up. We first identify categories of devices (phones, laptops, workstations, monitors, routers, etc.). For each device type, we research a range of common models, then use manufacturer data sheets or other available documentation to identify operational power draw (in watts) as well as embodied emissions related to manufacture and disposal (in kg CO$_2$e) per device. Based on this data, we model the power draw and embodied emissions of an ``exemplar'' device of each type.

For each device type we also estimate (1) a count (\eg, a laptop for every student and staff), (2) an estimated annual duty cycle (\eg, laptops used 8 hours per day, 200 days per year), and (3) a useful lifetime (in years).

Given all this, for each device type we can straightforwardly obtain an annual energy use (kWh) and annual embodied emissions (embodied emissions amortized over useful lifetime) (kg CO$_2$e).

We then estimate the local grid's emissions intensity (g CO$_2$e / kWh) to convert the energy use to emissions and obtain a combined total annual emissions number. We obtain emissions intensity data from Electricity Maps~\cite{electricitymaps}.

In addition to this simple linear mode, the model can also operate stochastically, instantiating devices from a distribution. However, we have no good data on the precise form of such a distribution, so we limit our analysis to the linear mode.

@@

We measure the impact of cloud servers in terms of annual operational emissions and amortized emissions due to manufacturing and disposal, in units of kg CO$_2$e / year.

We do not attempt to model application-level cloud services.

@@

\begin{description}
  \item[Cloud data center PUE] As of 2025, industry average power usage effectiveness is 1.56, with Google averaging 1.09 across its fleet~\cite{google:datacenter:efficiency} and Microsoft ranging from 1.12 to 1.30~\cite{microsoft:datacenter:efficiency}. We choose a PUE of 1.2 as representative of a state-of-the-art hyperscaler data center.
  \item[On-premise data center PUE] Based on limited historical power data, we assume an on-premise PUE of 2.0, attributable to inefficient cooling.
  \item[]  
\end{description}

@@

\subsection{Campus emissions}

@@

\begin{center}
\begin{tabular}{|r|l|}
\hline
\textbf{Production \& Disposal} & \\
\textbf{(kg CO$_2$e)} & \textbf{Source} \\ \hline
121 & \textcite{ecoinvent}  \\ \hline
104 & \textcite{teehan2013} (low) \\ \hline
338 & \textcite{teehan2013} (high) \\ \hline
244 & \textcite{rarecoil} \\ \hline
202 & \textcite{unctadder2024} \\ \hline
\end{tabular}
\label{tab:embodied_emissions:laptops}
\end{center}

The Ecoinvent numbers appear to be outliers on the low end.
The most comprehensive data comes from \textcite{rarecoil}, which lists manufacturer
information for over 90 popular devices. Mean supply chain emissions are
244 kg CO$_2$e with a standard deviation of 128 kg.
\assumption{We model laptop supply chain emissions using a truncated normal distribution with those characteristics.}

@@

\begin{center}
\begin{tabular}{|r|l|}
\hline
\textbf{Production \& Disposal} & \\
\textbf{(kg CO$_2$e)} & \textbf{Source} \\ \hline
238 & \textcite{ecoinvent} \\ \hline
303 & \textcite{teehan2013} \\ \hline
403 & \textcite{unctadder2024} \\ \hline
289 & \textcite{dellpcf} \\ \hline
277 & \textcite{boavizta} \\ \hline
\end{tabular}
\label{tab:embodied_emissions:desktops}
\end{center}

Most estimates are in the range of 250 - 400 kg CO$_2$e. The most detailed data comes from
Dell's Product Carbon Footprints (\cite{dellpcf}), which they publish for each of their products.
A sample of 15 desktops has mean supply chain emissions of 289 kg CO$_2$e with $\sigma = 80$ kg.
\assumption{We model desktop supply chain emissions using these statistics.}

@@

\begin{center}
\begin{tabular}{|r|l|}
\hline
\textbf{Production \& Disposal} & \\
\textbf{(kg CO$_2$e)} & \textbf{Source} \\ \hline
383 & \textcite{teehan2013} \\ \hline
1252 & \textcite{davy2021} (Dell) \\ \hline
1912 & \textcite{davy2021} (EC2) \\ \hline
899 & \textcite{boavizta} (medium server)\\ \hline
\end{tabular}
\label{tab:embodied_emissions:servers}
\end{center}

\textcite{davy2021} reports data for many Dell servers and estimates for different types of EC2 instances.
The former have mean supply chain emissions of 1252 kg CO$_2$e with $\sigma = 330$ kg, the latter
1912 kg with $\sigma = 885$ kg.
\assumption{We model server supply chain emissions using the statistics corresponding to Dell servers.}

@@

\begin{center}
\begin{tabular}{|r|l|}
\hline
\textbf{Production \& Disposal} & \\
\textbf{(kg CO$_2$e)} & \textbf{Source} \\ \hline
68 & \textcite{googlepixel8} (Pixel 8)\\ \hline
54 & \textcite{appleiphone13} (iPhone 13)\\ \hline
50 & \textcite{unctadder2024} \\ \hline
50 & \textcite{lovehagen2023} \\ \hline
\end{tabular}
\label{tab:embodied_emissions:phones}
\end{center}

\assumption{We model phone supply chain emissions using mean 50 kg CO$_2$e and $\sigma = 10$ kg.}

@@

\begin{description}
    \item[Computer monitors] 344 kg CO$_2$e (\textcite{teehan2013}, \textcite{dellpcf})
    \item[Conference room displays] 753 kg CO$_2$e (scaling monitor by $(40''/27'')^2$)
    \item[Printer/copier stations] 1167 kg CO$_2$e (\textcite{ecoinvent})
    \item[Network equipment] We did not find reliable lifecycle assessments for routers and switches,
      and have no information about UZH network architecture. The vast majority ($80-95\%$) of
      lifecycle emissions of network gear stem from usage (\textcite{cisco2024}, \textcite{jacob2023}).
      For now, we model network equipment as a 5\% overhead on servers (1 router or switch for 20 servers).
      (A typical data center rack contains 42 1U servers and one switch or router.)
\end{description}

@@

Lacking information about UZH's workloads, we model a hypothetical UZH cloud footprint as follows:
\begin{itemize}
    \item One-for-one replacement of one campus server with one Azure D8S\_v3 instance (2x Intel
      Xeon, 8 cores, 32 GB RAM, 2TB SSD) (\cite{msftvms}).
    \item Deployment to Microsoft Azure's ``\href{https://datacenters.microsoft.com/globe/explore?info=region_switzerlandnorth}{Switzerland North}'' region .
    \item Estimation of annual emissions via Datavizta, assuming constant 50\% load: 31 kg CO$_2$e for usage,
      97 kg CO$_2$e for manufacturing.
\end{itemize}

@@

Assuming a relatively clean grid at 200 g CO$_2$e / kWh (a low estimate), and that each server has embodied emissions of 4 t CO$_2$e (a high estimate),
we obtain:
\begin{itemize}
    \item 100 t CO$_2$e / day operational emissions.
    \item 260 t CO$_2$e training emissions.
    \item 16 kt CO$_2$e total embodied emissions.
\end{itemize}

After one year of operation, operational emissions are approximately 36 kt CO$_2$e, whereas training and embodied emissions are about 16 kt CO$_2$e,
so operational emissions account for $\approx 70\%$ of the total. With less conservative assumptions (dirtier grid, lower embodied emissions),
the proportion of operational emissions increases further.

@@

@@

Despite these limitations, our bottom-up model aligns closely with two UZH data points. First, the model correctly estimates approximately $400$ kW of compute for central campus data centers. Second, for university-owned campus IT infrastructure, the model estimates $1.4$ kt CO$_2$e emissions per year, close to the $1.45$ kt reported in the 2024 university sustainability report \cite{uzh:sustainability:report}.

@@

\item Under the assumption that the vast majority (90\%) of UZH students and staff have a laptop and smartphone, manufacturing personal devices accounts for over half of UZH's IT carbon footprint. For the estimated 32,000 laptops and smartphones, embodied emissions total $\sim 2400$ t CO$_2$e annually. Usage-related emissions amount to about a tenth of that.

@@

\item Embodied emissions are an important consideration for university-owned devices as well. If we exclude personal devices and LLMs, then total IT emissions shrink to $\sim 1500$ t CO$_2$e, of which $\frac{2}{3}$ comprises the embodied emissions of classroom computers, wall displays, printers, etc.

\item Data center cooling and efficiency represents a potential savings opportunity. The largest operational emissions are from data center servers and cooling (320 t). UZH data centers currently require as much energy for cooling as for computation, compared to the $6-15\%$ cooling overhead of cloud industry leaders. Moving workloads to efficient cloud providers could reduce data center emissions from $\sim 560$ t to $\sim 135$ t CO$_2$e annually, though this estimate requires further validation.

>>>>

\documentclass[11pt]{article}
\usepackage[margin=0.9in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{libertine}
\renewcommand{\sfdefault}{fvs} % Use as sans-serif
\renewcommand{\ttdefault}{fvm} % Use as monospace
\renewcommand{\ttfamily}{\small\fontfamily{\ttdefault}\selectfont} % Slightly smaller monospace font
\newcommand{\assumption}[1]{{#1}}
\newcommand{\eg}{{\em e.g.}}
\newcommand{\coe}{CO$_2$e}
\newcommand{\gcoe}{g \coe}
\newcommand{\kgcoe}{k\gcoe}
\newcommand{\gcoekwh}{\gcoe\ / kWh}
%\usepackage{helvet} % Helvetica font for sans-serif
%\usepackage{courier} % Courier font for monospace
\usepackage{hyperref}
\hypersetup{
    hidelinks
}
\usepackage[style=numeric]{biblatex}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{titling}

@@

Information technology accounts for over 5\% of global electricity consumption and at least 1.7\% of global carbon emissions~\cite{wb:itu:ict}. At the University of Zürich (UZH), which with over 38,000 students and staff is the largest Swiss university, IT in 2024 accounted for about 1.5 Mt \coe\ (megatons of carbon dioxide equivalent) of greenhouse gas (GHG) emissions, approximately 6.3\% of total university emissions in 2024~\cite{uzh:sustainability:report}. The cited report takes into account university-owned workstations, servers, and audiovisual equipment, but does include other IT-related activities such as networking, cloud computing, or the growing use of artificial intelligence, and in particular the adoption of generative AI and large language models (LLMs).

@@

\subsubsection*{On-campus devices}

@@

University-owned devices are in scope. We have accurate numbers for staff workstations, classroom computers, network access points, and audiovisual devices managed by the central IT department. We can only estimate other devices, such as networking equipment or computers managed by individual research groups. As before, we include these devices' embodied emissions in the model.

@@

Due to lack of data, we do not attempt to model emissions from application-level cloud services such as Microsoft 365 or Google Workspace. This omission leads to undercounting total IT energy use but to overestimating the potential relative impact of AI.

UZH faculties do make use of shared off-campus supercomputing facilities, primarily the Swiss National Supercomputing Center, CSCS~\cite{cscs}. We estimate the operational impact of these workloads, but, again due to lack of data, we do not include embodied emissions. Again, this undercounts total university IT energy use but overestimates the potential relative impact of AI.

@@

We estimate inference-time emissions of current large language models based on the literature. We also include training-time emissions and embodied emissions of language models (thouse these are harder to estimate and we err on the side of overestimating them).

@@

\subsection{Accounting approach}

\subsubsection*{On-campus devices}

We model campus device emissions bottom-up. We first identify categories of devices (phones, laptops, workstations, monitors, routers, etc.). For each device type, we research a range of common models, then use manufacturer data sheets or other available documentation to identify operational power draw (in watts) as well as embodied emissions related to manufacture and disposal (in \kgcoe) per device. Based on this data, we model the power draw and embodied emissions of an ``exemplar'' device of each type.

For each device type we also estimate (1) a count (\eg, a laptop for every student and staff), (2) an estimated duty cycle (\eg, laptops used 8 hours per day, 200 days per year), and (3) a useful lifetime (in years).

Given all this, for each device type we can straightforwardly obtain an annual energy use (kWh) and annual embodied emissions (embodied emissions amortized over useful lifetime) (\kgcoe).

We then estimate the local grid's emissions intensity (\gcoekwh) to convert the energy use to emissions and obtain a combined total annual emissions number. We obtain emissions intensity data from Electricity Maps~\cite{electricitymaps}.

In addition to this simple linear mode, the model can also operate stochastically, instantiating devices from a distribution. However, we have insufficient data on the precise form of such a distribution, so we limit our analysis to the linear mode.

@@

We measure the impact of cloud servers in terms of annual operational emissions and amortized emissions due to manufacturing and disposal, in units of \kgcoe\ / year.

As mentioned earlier, we do not attempt to model application-level cloud services.

@@

\begin{description}
  \item[Cloud data center PUE] As of 2025, industry average power usage effectiveness is 1.56, with Google averaging 1.09 across its fleet~\cite{google:datacenter:efficiency} and Microsoft ranging from 1.12 to 1.30~\cite{microsoft:datacenter:efficiency}. We choose a PUE of 1.2 as representative of a state-of-the-art hyperscaler data center.
  \item[On-premise data center PUE] Based on limited UZH historical power data, we assume an on-premise PUE of 2.0, attributable to inefficient cooling.
  \item[Device duty cycle] We make unsupported but plausible assumptions about the number of days that different devices are used (\eg, end-user devices are not used every day) and their duty cycle throughut the day.
  \item[Electricity emissions intensity for on-premise data centers] For Switzerland, we can compute average hourly emissions intensity over 2024 based on data from Electricity Maps~\cite{electricitymaps}. Emissions vary from a high of 64 \gcoekwh\ in the middle of the night to a low of 45 \gcoekwh\ in the middle of the day. This variation is sufficiently large that we refine the model's device duty cycle to four six-hour periods over the course of the day. It allows us to specify, for example, that user-devices are mostly used during the day whereas servers run around the clock.
  \item[Electricity emissions intensity for cloud data centers and LLMs] These facilities are unlikely to be located in Switzerland, and Switzerland's grid is cleaner than average. We use the European Union's average emissions intensity of electricity generation in 2023, 207 \gcoekwh~\cite{eea:emissions:2025}.
  \item[LLM usage] We must make assumptions about the fraction of the university population that use LLMs and the frequency of use (queries / day).
\end{description}

@@

The previous section describes our overall modeling approach, which we believe is broadly applicable to university and other large campus settings. This section presents the reasoning behind specific default values we employ in the UZH evaluation.

\subsection{On-campus devices}

\subsubsection*{Number of devices}

We have an authoritative census of devices (workstations, servers, audiovisual equipment) managed by central IT, as follows.

\begin{center}
  \begin{tabular}{|l|r|r|}
  \hline
  \textbf{Device type} & \textbf{Count} & \textbf{Life (y)}\\ \hline
  Laptops & 10,440 & 4 \\
  Monitors & 10,400 & 6 \\
  Servers & 1,121 & 5 \\
  Printers & 333 & 6 \\
  Mobile phones & 2,439 & 3 \\
  A/V devices & 431 & 6 \\ \hline
  \end{tabular}
  \label{tab:oncampusdevices}
\end{center}

On top of this, we assume that

\subsubsection*{Operational emissions}

\subsubsection*{Embodied emissions}

We obtain

\subsection{Cloud and supercomputing facilities}

\subsection{LLMs}

@@

\begin{center}
\begin{tabular}{|r|l|}
\hline
\textbf{Production \& Disposal} & \\
\textbf{(\kgcoe)} & \textbf{Source} \\ \hline
121 & \textcite{ecoinvent}  \\ \hline
104 & \textcite{teehan2013} (low) \\ \hline
338 & \textcite{teehan2013} (high) \\ \hline
244 & \textcite{rarecoil} \\ \hline
202 & \textcite{unctadder2024} \\ \hline
\end{tabular}
\label{tab:embodied_emissions:laptops}
\end{center}

The Ecoinvent numbers appear to be outliers on the low end.
The most comprehensive data comes from \textcite{rarecoil}, which lists manufacturer
information for over 90 popular devices. Mean supply chain emissions are
244 \kgcoe\ with a standard deviation of 128 kg.
\assumption{We model laptop supply chain emissions using a truncated normal distribution with those characteristics.}

@@

\begin{center}
\begin{tabular}{|r|l|}
\hline
\textbf{Production \& Disposal} & \\
\textbf{(\kgcoe)} & \textbf{Source} \\ \hline
238 & \textcite{ecoinvent} \\ \hline
303 & \textcite{teehan2013} \\ \hline
403 & \textcite{unctadder2024} \\ \hline
289 & \textcite{dellpcf} \\ \hline
277 & \textcite{boavizta} \\ \hline
\end{tabular}
\label{tab:embodied_emissions:desktops}
\end{center}

Most estimates are in the range of 250 - 400 \kgcoe. The most detailed data comes from
Dell's Product Carbon Footprints (\cite{dellpcf}), which they publish for each of their products.
A sample of 15 desktops has mean supply chain emissions of 289 \kgcoe\ with $\sigma = 80$ kg.
\assumption{We model desktop supply chain emissions using these statistics.}

@@

\begin{center}
\begin{tabular}{|r|l|}
\hline
\textbf{Production \& Disposal} & \\
\textbf{(\kgcoe)} & \textbf{Source} \\ \hline
383 & \textcite{teehan2013} \\ \hline
1252 & \textcite{davy2021} (Dell) \\ \hline
1912 & \textcite{davy2021} (EC2) \\ \hline
899 & \textcite{boavizta} (medium server)\\ \hline
\end{tabular}
\label{tab:embodied_emissions:servers}
\end{center}

\textcite{davy2021} reports data for many Dell servers and estimates for different types of EC2 instances.
The former have mean supply chain emissions of 1252 \kgcoe\ with $\sigma = 330$ kg, the latter
1912 kg with $\sigma = 885$ kg.
\assumption{We model server supply chain emissions using the statistics corresponding to Dell servers.}

@@

\begin{center}
\begin{tabular}{|r|l|}
\hline
\textbf{Production \& Disposal} & \\
\textbf{(\kgcoe)} & \textbf{Source} \\ \hline
68 & \textcite{googlepixel8} (Pixel 8)\\ \hline
54 & \textcite{appleiphone13} (iPhone 13)\\ \hline
50 & \textcite{unctadder2024} \\ \hline
50 & \textcite{lovehagen2023} \\ \hline
\end{tabular}
\label{tab:embodied_emissions:phones}
\end{center}

\assumption{We model phone supply chain emissions using mean 50 \kgcoe\ and $\sigma = 10$ kg.}

@@

\begin{description}
    \item[Computer monitors] 344 \kgcoe\ (\textcite{teehan2013}, \textcite{dellpcf})
    \item[Conference room displays] 753 \kgcoe\ (scaling monitor by $(40''/27'')^2$)
    \item[Printer/copier stations] 1167 \kgcoe\ (\textcite{ecoinvent})
    \item[Network equipment] We did not find reliable lifecycle assessments for routers and switches,
      and have no information about UZH network architecture. The vast majority ($80-95\%$) of
      lifecycle emissions of network gear stem from usage (\textcite{cisco2024}, \textcite{jacob2023}).
      For now, we model network equipment as a 5\% overhead on servers (1 router or switch for 20 servers).
      (A typical data center rack contains 42 1U servers and one switch or router.)
\end{description}

@@

Lacking information about UZH's workloads, we model a hypothetical UZH cloud footprint as follows:
\begin{itemize}
    \item One-for-one replacement of one campus server with one Azure D8S\_v3 instance (2x Intel
      Xeon, 8 cores, 32 GB RAM, 2TB SSD) (\cite{msftvms}).
    \item Deployment to Microsoft Azure's ``\href{https://datacenters.microsoft.com/globe/explore?info=region_switzerlandnorth}{Switzerland North}'' region .
    \item Estimation of annual emissions via Datavizta, assuming constant 50\% load: 31 \kgcoe\ for usage,
      97 \kgcoe\ for manufacturing.
\end{itemize}

@@

Assuming a relatively clean grid at 200 \gcoekwh (a low estimate), and that each server has embodied emissions of 4 t \coe\ (a high estimate),
we obtain:
\begin{itemize}
    \item 100 t \coe\ / day operational emissions.
    \item 260 t \coe\ training emissions.
    \item 16 kt \coe\ total embodied emissions.
\end{itemize}

After one year of operation, operational emissions are approximately 36 kt \coe, whereas training and embodied emissions are about 16 kt \coe,
so operational emissions account for $\approx 70\%$ of the total. With less conservative assumptions (dirtier grid, lower embodied emissions),
the proportion of operational emissions increases further.

@@

\subsection{Model validation}

Using Ecoinvent embodied emissions values where available, and only the numbers and types of devices reported in the UZH 2024 Sustainability Report~\cite{uzh:sustainability:report}, our model estimates emissions of 1,460 t \coe, within 1\% of the 1,454 estimated in the UZH report, even though the methodology is somewhat different. Our model exposes average device power ratings and duty cycle, grid emissions intensity, etc., directly, whereas the approach used in the sustainability report relies on gross emissions factors (\kgcoe/device for embodied emissions, \kgcoe/device/day for operational emissions) provided by Ecoinvent.

While this is admittedly just one data point, it provides some confidence that our assumptions are realistic and that the model can be used to extrapolate emissions and trends beyond the scope of the sustainability report.

@@

Despite these limitations, our bottom-up model aligns closely with two UZH data points. First, the model correctly estimates approximately $400$ kW of compute for central campus data centers. Second, for university-owned campus IT infrastructure, the model estimates $1.4$ kt \coe\ emissions per year, close to the $1.45$ kt reported in the 2024 university sustainability report \cite{uzh:sustainability:report}.

@@

\item Under the assumption that the vast majority (90\%) of UZH students and staff have a laptop and smartphone, manufacturing personal devices accounts for over half of UZH's IT carbon footprint. For the estimated 32,000 laptops and smartphones, embodied emissions total $\sim 2400$ t \coe\ annually. Usage-related emissions amount to about a tenth of that.

@@

\item Embodied emissions are an important consideration for university-owned devices as well. If we exclude personal devices and LLMs, then total IT emissions shrink to $\sim 1500$ t \coe, of which $\frac{2}{3}$ comprises the embodied emissions of classroom computers, wall displays, printers, etc.

\item Data center cooling and efficiency represents a potential savings opportunity. The largest operational emissions are from data center servers and cooling (320 t). UZH data centers currently require as much energy for cooling as for computation, compared to the $6-15\%$ cooling overhead of cloud industry leaders. Moving workloads to efficient cloud providers could reduce data center emissions from $\sim 560$ t to $\sim 135$ t \coe\ annually, though this estimate requires further validation.
